---
title: "E-Commerce Sales and Customer Intelligence Dashboard"
author: "Mohammad Qasim"
date: "2026-10-05"
output: pdf_document
---
\newpage

# E-Commerce Business Intelligence Dashboard as a real portfolio project.

We’ll do it in stages:

\begin{enumerate}

\item{Define the business problem and requirements}
\item{Create the e-commerce dataset/database}
\item{Load data into MySQL}
\item{Perform advanced SQL analysis}
\item{Clean/transform data with Power Query}
\item{Build a proper Star Schema}
\item{Create DAX measures}
\item{Build the Power BI dashboard}
\item{Extract business insights and recommendations}
\end{enumerate}

## Business scenario
Me as a Data Analyst for an e-commerce company and
the management team wants answers to questions such as:
\begin{itemize}
\item{How much revenue are we generating?}
\item{Which products and categories perform best?}
\item{Which customers generate the most revenue?}
\item{Which regions are performing well?}
\item{Are sales increasing or decreasing?}
\item{What is our profit margin?}
\item{What is the average order value?}
\item{How many customers are returning?}
\item{Which products have poor profitability?}
\item{How is performance changing month-over-month?}
\end{itemize}

## Our technology stack
We'll deliberately use the technologies you want to strengthen:
                  
So by the end, I will be able to say
`I built an end-to-end BI solution starting from relational data in MySQL, performed advanced SQL analysis, transformed the data using Power Query, created a star-schema model, developed DAX measures and built an interactive Power BI dashboard.`

## Business requirements
 his is an important part that most student Power BI projects skip.

We'll create a small Business Requirements Document.

*Stakeholders*

*E-Commerce Management*

### Business objective:
 Develop an interactive business intelligence dashboard that enables management to monitor sales, profitability, customer behavior and product performance.
 
## Key business questions:

\begin{itemize}
\item{Sales}
\begin{enumerate}
\item{What is total revenue?}
\item{What is total profit?}
\item{What is the monthly sales trend?}
\item{What is the month-over-month growth?}
\end{enumerate}

\item{Customers}
\begin{enumerate}
\item{Who are our highest-value customers?}
\item{What percentage of customers are returning?}
\item{Which customer segments generate the most revenue?}
\end{enumerate}

\item{Products}
\begin{enumerate}
\item{Which products generate the most revenue?}
\item{Which products generate the most profit?}
\item{Which categories have poor margins?}
\end{enumerate}

\item{Geography}
\begin{enumerate}
\item{Which regions generate the most revenue?}
\item{Which stores perform best?}
\end{enumerate}

\item{Management}
\begin{enumerate}
\item{What are our key KPIs?}
\item{Where are the biggest opportunities for improvement?}
\end{enumerate}
\end{itemize}

## Define our KPIs:

| KPI                     | Meaning                         |
|-------------------------|---------------------------------|
| **Total Revenue**       | Total sales generated           |
| **Total Profit**        | Revenue - cost                  |
| **Profit Margin**       | Profit as % of revenue          |
| **Total Orders**        | Number of orders                |
| **Total Customers**     | Unique customers                |
| **Units Sold**          | Total quantity sold             |
| **Average Order Value** | Revenue / orders                |
| **MoM Growth**          | Month-over-month revenue growth |


We're going to build the underlying data first.

## Dataset
For this project, I recommend that I create our own realistic dataset rather than downloading a random dataset.

**That gives us control over:** 

 - SQL relationships 
 - Data quality issues 
 - Missing values 
 - Duplicate records 
 - Dates
 - Customers
 - Products
 - Stores
 - Revenue
 - Cost
 - Discounts

And it lets us deliberately create a dataset that demonstrates the skills employers are looking for.

### Target size:

| Category   | Count    |
|------------|---------:|
| Customers  | 1,000    |
| Products   | 100      |
| Stores     | 15       |
| Orders     | 10,000+  |
| Sales rows | 10,000+  |

Enough data to make the dashboard realistic without making it unnecessarily difficult to work with.

# First task:
I Should Required:

 - MySQL
 - MySQL Workbench
 - Power BI Desktop
 
**Recommended**

 - R/Python
 - Git
 - GitHub

I already have MySQL, Power BI, Python and R, so we're going to make those skills demonstrable through the project rather than just listing them.

# Results:

##  Sales

**Q: What is total revenue?**

```{r,echo=FALSE,}
t1=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Sales/Q1.csv")
require(knitr)
kable(t1)
```

**Q: What is total profit?**

```{r,echo=FALSE}
t2=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Sales/Q2.csv")

kable(t2)
```


**Q: What is the monthly sales trend?**

```{r,echo=FALSE}
t3=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Sales/Q3.csv")

kable(t3)
```


**Q:What is the month-over-month growth**

```{r,echo = FALSE}
t4=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Sales/Q4.csv")

kable(t4)
```

## Customers

**Q: Who are our highest-value customers?**
```{r,echo = FALSE}
t5=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Customers/Q1.csv")

kable(t5)
```

**Q: What percentage of customers are returning?**
```{r,echo = FALSE}
t6=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Customers/Q2.csv")

kable(t6)
```

**Q: Which customer segments generate the most revenue?**
```{r,echo=FALSE}
t7=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Customers/Q3.csv")

kable(t7)
```


## Products

**Q: Which products generate the most revenue?**
```{r,echo=FALSE}
t8=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Products/Q1.csv")

kable(t8)
```
**Q: Which products generate the most profit?**
```{r,echo=FALSE}
t9=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Products/Q2.csv")

kable(t9)
```

**Q: Which categories have poor margins?**
```{r,echo=FALSE}
t10=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Products/Q3.csv")

kable(t10)
```

## Geography

**Q: Which regions generate the most revenue?**
```{r}
t11=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Region/Q1.csv")

kable(t11)
```


**Q: Which stores perform best?**
```{r}
t12=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Region/Q2.csv")

kable(t12)
```

## Management 

**Q: What are our key KPIs?**
```{r}
t13=read.csv("C:/Users/M.QasimSalman/Desktop/MSc/MYSQL/ecommerce/Answers/Management/Q1.csv")

kable(t13)
```
**Q: Where are the biggest opportunities for improvement?**

